---
title: "Comparision of Risk in Investment in crypto and metaverse token"
author: "Ayan Bashir Sheikh"
date: "`r Sys.Date()`"
output: html_document
---




```{r}
suppressWarnings({
library(readxl)
d1=read_xlsx("D:\\ST 402 Project\\data\\metaverse tokens\\four tokens\\Axie Infinity.xlsx")
d2=read_xlsx("D:\\ST 402 Project\\data\\metaverse tokens\\four tokens\\Decentraland.xlsx")
d3=read_xlsx("D:\\ST 402 Project\\data\\metaverse tokens\\four tokens\\Enjin.xlsx")
d4=read_xlsx("D:\\ST 402 Project\\data\\metaverse tokens\\four tokens\\Sandbox.xlsx")
BTC = read_excel("E:/Bitcoin.xlsx")
ETH=read.csv("E:\\Ethereum Historical Data.csv")

#colnames(ETH)=colnames(BTC)=c("currency","Date","Close","Open","High","Low")
ETH$returns = c(0, diff(ETH$Price) / head(ETH$Price, -1))

})
```






```{r}
suppressWarnings({
library(PerformanceAnalytics)
VaR_95E = VaR(ETH$returns, p = 0.95, method = "historical");VaR_95E
CVaR_95E = ES(ETH$returns, p = 0.95, method = "historical");CVaR_95E
VaR_95B = VaR(BTC$Return, p = 0.95, method = "historical");VaR_95B
CVaR_95B = ES(BTC$Return, p = 0.95, method = "historical");CVaR_95B

VaR_95A = VaR(d1$Return, p = 0.95, method = "historical");VaR_95A
CVaR_95A = ES(d1$Return, p = 0.95, method = "historical");CVaR_95A
VaR_95D = VaR(d2$Return, p = 0.95, method = "historical");VaR_95D
CVaR_95D = ES(d2$Return, p = 0.95, method = "historical");CVaR_95D
VaR_95En = VaR(d3$Return, p = 0.95, method = "historical");VaR_95En
CVaR_95En= ES(d3$Return, p = 0.95, method = "historical");CVaR_95En
VaR_95s= VaR(d4$Return, p = 0.95, method = "historical");VaR_95s
CVaR_95s = ES(d4$Return, p = 0.95, method = "historical");CVaR_95s


risk_measures = data.frame(
  Token = c("ETH", "BTC", "LTC", "Axie infinity", "Decentraland", "Enjin", "Sandbox"),
  VaR = c(VaR_95E, VaR_95B, VaR_95L, VaR_95A, VaR_95D, VaR_95En, VaR_95s),
  CVaR = c(CVaR_95E, CVaR_95B, CVaR_95L, CVaR_95A, CVaR_95D, CVaR_95En, CVaR_95s)
)


risk_measures


})
```




**Ethereum` :** The historical VaR is -0.06859660 and the CVaR is -0.10896281. This means that, based on past data, the maximum loss expected from ETH in a day is 6.86% under normal market conditions. In the worst-case scenario (i.e., the worst 5% of cases), the expected loss is 10.90%.


**Bitcoin:** The historical VaR is -0.05342919 and the CVaR is -0.08864855. This means that, based on past data, the maximum loss expected from BTC in a day is 5.34% under normal market conditions. In the worst-case scenario, the expected loss is 8.86%.


**Litcoine:** The historical VaR is -0.07514402 and the CVaR is -0.11226267. This means that, based on past data, the maximum loss expected from LTC in a day is 7.51% under normal market conditions. In the worst-case scenario, the expected loss is 11.23%.


**Axie Infinity:** The historical VaR is -0.12117114 and the CVaR is -0.16534219. This means that, based on past data, the maximum loss expected from Axie Infinity in a day is 12.12% under normal market conditions. In the worst-case scenario, the expected loss is 16.53%.


**Decentraland:** The historical VaR is -0.11431650 and the CVaR is -0.17017849. This means that, based on past data, the maximum loss expected from Decentraland in a day is 11.43% under normal market conditions. In the worst-case scenario, the expected loss is 17.02%.


**Enjin:** The historical VaR is -0.10918033 and the CVaR is -0.16748954. This means that, based on past data, the maximum loss expected from Enjin in a day is 10.92% under normal market conditions. In the worst-case scenario, the expected loss is 16.75%.

**Sandbox:** The historical VaR is -0.12068350 and the CVaR is -0.17679551. This means that, based on past data, the maximum loss expected from Sandbox in a day is 12.07% under normal market conditions. In the worst-case scenario, the expected loss is 17.68%.


```{r}

df=data.frame(d1$Return,d2$Return,d3$Return,d4$Return,BTC$Return,ETH$returns)

```






```{r}
# Load necessary libraries
library(PerformanceAnalytics)
library(rugarch)

# Assuming 'df' is your data frame and 'returns' is the column of returns
# Replace 'df' and 'returns' with your actual data frame and returns column

# Calculate VaR and CVaR for each cryptocurrency
colnames(df)= c("axie_infinity_returns", "decentraland_returns", "sandbox_returns", "enjin_returns", "bitcoin_returns", "ethereum_returns")
  cryptos <- c("axie_infinity", "decentraland", "sandbox", "enjin", "bitcoin", "ethereum")
for (crypto in cryptos) {
  returns <- df[[paste0(crypto, "_returns")]]
  
  # Fit a GARCH(1,1) model
  spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)), 
                     mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), 
                     distribution.model = "norm")
  fit <- ugarchfit(spec, data = returns)
  
  # Forecast the next day volatility
  forecast <- ugarchforecast(fit, n.ahead = 1)
  sigma <- as.numeric(forecast@forecast$SigmaFor)
  
  # Calculate VaR and CVaR
  VaR_95 <- qnorm(0.05, mean = 0, sd = sigma)
  CVaR_95 <- (1 / (1 - 0.95)) * dnorm(qnorm(0.95)) * sigma
  
  # Print the results
  print(paste0("VaR at 95% for ", crypto, ": ", VaR_95))
  print(paste0("CVaR at 95% for ", crypto, ": ", CVaR_95))
}


```




```{r}
VaRs = c(VaR_95E, VaR_95B, VaR_95A, VaR_95D, VaR_95En, VaR_95s)
  CVaRs = c(CVaR_95E, CVaR_95B, CVaR_95A, CVaR_95D, CVaR_95En, CVaR_95s)
cryptos <- c("axie_infinity", "decentraland", "sandbox", "enjin", "bitcoin", "ethereum")
data <- data.frame(cryptos, VaRs, CVaRs)

# Plot
library(ggplot2)
ggplot(data, aes(x=cryptos, y=VaRs)) + 
  geom_bar(stat="identity", fill="lightblue") +
  geom_point(aes(y=CVaRs), color="red", size=3) +
  theme_minimal() +
  labs(title="VaR and CVaR for Cryptocurrencies", x="Cryptocurrency", y="Value") +
  geom_text(aes(label=VaRs, y=VaRs), vjust=1.5, color="white") +
  geom_text(aes(label=CVaRs, y=CVaRs), vjust=-0.5, color="red")

# Table
library(knitr)
kable(data, caption = "VaR and CVaR for Cryptocurrencies")

```


```{r}

token_VaRs = c( VaR_95A, VaR_95D, VaR_95En, VaR_95s)
  token_CVaRs = c( CVaR_95L, CVaR_95A, CVaR_95D, CVaR_95En, CVaR_95s)
  
crypto_VaRs=c(VaR_95E,VaR_95B)
crypto_CVaRs=c(CVaR_95E,CVaR_95B)

# Wilcoxon Rank Sum test to compare VaR between cryptocurrencies and Metaverse tokens
wilcox_test_VaR = wilcox.test(crypto_VaRs, token_VaRs)
print(paste("p-value for the difference in VaR: ", wilcox_test_VaR$p.value))

# Wilcoxon Rank Sum test to compare CVaR between cryptocurrencies and Metaverse tokens
wilcox_test_CVaR= wilcox.test(crypto_CVaRs, token_CVaRs)
print(paste("p-value for the difference in CVaR: ", wilcox_test_CVaR$p.value))

  

```


```{r}
data <- risk_measures

# Melt the data frame to long format for easy plotting with ggplot2
data_long <- melt(data, id.vars = 'Token')

# Create a time series plot
ggplot(data_long, aes(x = Token, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "VaR and CVaR for Cryptocurrencies and Tokens", x = "Cryptocurrency and Tokens", y = "Value", fill = "Measure")

```